Named Entity Recognition on HTML as a web data extraction subtask

### 2018-09-21 tex/article.tex
Final version for article.

\usepackage[latin1]{inputenc}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{amsfonts}
\usepackage{multirow}

@@

Reliable researcher affiliation data is necessary to allow enquiring
about international research group productivity and publication patterns.
Public bibliographic databases such as DBLP and Google Scholar hold
invaluable data about the academic environment. However, the researcher
affiliation information is frequently missing or outdated.
We propose a statistical data extraction method to acquire affiliation 
information directly from university websites and solve the name extraction
task in general.
Previous approaches to web data extraction either lack in flexibility,
because wrappers do not generalize well on cross website tasks, or 
they lack in precision, because domain agnostic methods neglect 
useful properties of this particular application domain.
Our statistical approach solves the name extraction task with 
a framework that incorporates both textual and structural features to
yield an outstanding tradeoff between generality and precision. 
We conducted experiments over a collection of 152 faculty 
web pages in multiple languages from universities in 49 countries
and obtained 94.40\% precision, 97.61\% recall and 0.9597 F-measure at
the extraction task.

@@

% ==========================================
% Beginning of text.                       |
% ==========================================

@@

HTML documents most often lie in between the structured / unstructured data paradigm.
DOM hierarchy, element disposition, class names, and other features related to 
the document structure and indirectly associated with the data itself can be valuable 
information in the task of identifying entities and determining relationships. Yet we 
cannot expect these features to be completely constrained by an underlying pattern. 
Organization patterns tend to follow some guidelines but are in no way subject to 
strict rules. That is why classical information extraction systems such as automatic
wrapper generators usually do not translate well across different websites.

@@

In recent years, we saw an astonishing progress in Natural Language Processing.
However, despite WDE being a closely related field of research,
many tools rely on heuristics or hand coded rules to perform record segmentation,
and many methods resort to regular expressions or gazetteer matching to perform attribute
labelling. While these approaches can be sufficient to extract information from webpages
with similar templates, and identifying attributes with distinct formats like prices and
dates, we still lack a sufficiently flexible and intelligent approach to handle more complicated cases.
Even when extracting data from web documents only, data records may contain pure text with 
relevant named entities. Also, many WDE algorithms can effectively exploit the semi-
structure of web documents, but when a method relies too much on structural features, it
often produces poor generalization on cross website extraction tasks. Contrastingly, we 
understand that statistical based approaches can be much more flexible.

@@

In 2008, Sarawagi \cite{Sarawagi2008} produced a survey that classified wrappers into
rule-based methods, statistical methods and hybrid models, bringing together 
the fields of named entity recognition, relationship extraction and information extraction. 
The rule based methods encompass most of the 
previous models. The statistical methods convert the extraction task into a token labeling 
task, identifying the target entities through the assignment of labels as in a typical 
Named Entity Recognition task. Traditionally, Hidden Markov Models \cite{Leek1997, Freitag1999}, 
Linear Chain Conditional Random Fields \cite{Lafferty2001}, and Maximum Entropy Taggers 
\cite{McCallum2000} have been the usual choice for linear sequence tagging models.
More recently, with the advancement in Natural Language Processing and Deep Learning, 
neural models outperformed previous NER methods. Huang \cite{Huang2015} introduced the 
bidirectional Long Short-Term Memory (LSTM) model with a Conditional Random Field (CRF) output layer
for NER. Lample \cite{Lample2016} incorporated Convolutional Neural Network based character representations 
on top of the architecture. And Ma \cite{Ma2016} introduced
LSTM based character representations. Both obtaining improvements with systems that are less reliant
on feature engineering.

@@

Most data extraction systems focus on extracting information from single websites
and are therefore unsuited for cross website extraction tasks. Even unsupervised
approaches that are domain independent, such as RoadRunner \cite{Crescenzi2001} 
and EXALG \cite{Arasu2003} only work well for extracting data from pages generated 
from the same template. A statistical approach to unsupervised domain 
independent web data extraction was described by Zhu et al \cite{Zhu2005}. The 2D CRF 
model takes a web page segmented into data blocks and employs a two dimensional conditional 
random field model to perform attribute labelling. The model was further improved in 
\cite{Zhu2006} to model record segmentation and attribute labeling as a joint task.
Some of the limitations of early unsupervised methods 
were also tackled by ObjectRunner (\cite{Abdessalem2010}) and AMBER (\cite{Furche2012}). 
These methods work by annotating webpages automatically with regular expressions, gazetteers and 
knowledge bases. They can rectify low quality annotations and even improve the annotators
by exploring regular structures in the DOM during the record segmentation phase.

@@

Most web data extraction systems rely on hand crafted rules or gazetteers to perform
attribute annotation. Machine learning approaches to NER can improve annotations of 
more complex entities and even manage entity detection without resorting to a gazetteer.
We explored multiple approaches to the Named Entity Recognition problem in the context 
of a web information extraction task. We first introduce two traditional approaches, 
namely: Hidden Markov Models and Linear Chain Conditional Random Fields. And further
we explore neural architectures.

@@

A Markov Model is a stochastic model that computes the most probable sequence of states 
given a limited set of observable states $ \S = \{s_1, s_2, ..., s_n \} $.
The Hidden Markov Model (HMM) differs from the Markov Model in that it
does not observe the states directly, but rather a probabilistic function of those 
states. For example in NER, the words are observed, however the Named Entity labels
associated with these words are not. Formally, we want to compute the most probable
sequence of labels $ Y = \{y_1, y_2, ..., y_n\} $ for a sequence of observed tokens
$ X = \{x_1, x_2, ..., x_n\} $.

\begin{equation}
Y* = \argmax_{Y} P(Y|X)
\end{equation}

@@

Since $ P(X) $ is the same for all label sequences $ Y $, we can simply maximize:

\begin{equation}
P(Y|X) \propto P(X|Y) P(Y)
\end{equation}

@@

HMM based taggers have been succesfully apllied in many NLP and IE tasks 
\cite{Rabiner1989, Leek1997, Freitag2000}. They are incredibly fast to train and very 
interpretable, making them a good choice for a first approximation. However, these models 
are highly dependent on the right selection of features, what may outweigh the benefit of a 
small training cost.

@@

A Linear Chain Conditional Random Field (CRF) is the discriminative analog to the HMM,
it was first introduced by \cite{Lafferty2001}. It is a distribution $ P(Y|X) $ that takes the form:

\begin{equation}
P(Y|X) = \frac{1}{Z(x)} \prod_{t=1}^{T} exp \left( \sum_{k=1}^{K} \theta_k f_k(y_{t-1}, y_t, X) \right)
\end{equation}

@@

\begin{equation}
Z(x) = \sum_{Y} \prod_{t=1}^{T} exp \left( \sum_{k=1}^{K} \theta_k f_k(y_{t-1}, y_t, x) \right)
\end{equation}

which is a sum over all possible label assignments $ Y $. The partition function can be efficiently
and exactly calculated with the sum-product algorithm. Parameter estimation is usually done through 
negative log likelihood minimization. The function can be optimized with techniques suitable for other 
maximum entropy models such as L-BFGS \cite{Liu1989}. The most likely label sequences can be decoded 
with the Viterbi algorithm like in the case of HMMs.

@@

\subsection{Neural Models}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{pics/cnn}
  \caption{CNN}
  \label{fig:cnn}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{pics/lstm_char_representations}
  \caption{LSTM character representations}
  \label{fig:lstm_char}
\end{figure}

@@

\section{NER on HTML Dataset}

@@

This task would be similar to the task of labeling authors in comments
However, we wanted to evaluate models without relying on any sophisticated data record
segmentation system. In many cases, entity annotation may precede the segmentation
phase in web data extraction methods, so annotators have to be able to work with
raw HTML data. Because of the type of documents present in the dataset, many sentences 
contained only one or two tokens, while other documents contained long texts.
Finally, all tokens were tagged using the IOB scheme put forward by
Ramshaw and Marcus \cite{Ramshaw1995}.

@@

\begin{table}[h]
  \small
  \begin{center}
    \begin{tabular}{ lllll }
      \toprule
      Data file & Precision & Recall & F1 & Correct names \\
      \midrule
      Training   & 0.7316 & 0.2303 & 0.3504 & 1341 from 5822 \\ 
      Validation & 0.8474 & 0.2858 & 0.4274 & 511 from 1788 \\ 
      Test       & 0.8717 & 0.3287 & 0.4773 & 890 from 2708 \\ 
      \bottomrule
    \end{tabular}
  \end{center}
  \caption{Gazetteer coverage in each data file}
  \label{tab:gazetteer}
\end{table}

@@

We conducted two experiments to evaluate the best models for named entity recognition
on HTML in the context of web data extraction. In the first experiment, we used the
whole training set (90 documents). In the second experiment, the training set was split 
into five non overlapping sets with only 18 documents, but the test set was kept unchanged. 
The goal of this second experiment 
was to evaluate models when training data is scarce, since labelling data can be a very 
time consuming task.

Evaluation of model performance in this task was measured with the precision, recall and 
F1 scores (Van Rijsbergen, 1975). Precision is the percentage of named entities found by 
the model that are correct. Recall is the percentage of named entities that are present
in the corpus and were found by the model. The F1 score is a composite measure that combines
precision and recall with the formula:

@@

Named entities were only considered to be correct if they were a complete match of the 
corresponding entity in the dataset. The models evaluated are presented in Table \ref{tab:models}.
We evaluated the models using the complete set of features, presented in 
table \ref{tab:features} and no features besides word embeddings for neural models and
CRFs, and unaccented words for HMMs. To represent models with the complete set of features
we added a "+F" suffix to the model name. Additionally, HMMs were evaluated with and
without the self training strategy. To represent HMMs that use the self training strategy
we added the suffix "+ST" to the model name.
Results were reported for the validation and test sets in experiment 1 and only for the
test set in experiment 2. In the second experiment, precision, recall and F1 scores were 
averaged over the five runs. The standard deviation is presented next to the measures.

\subsection{Experiment 1: complete dataset}

\begin{table}[h]
  \small
  \begin{center}
    \begin{tabular}{ lllllll }
      \toprule
      \multirow{2}{*}{Model} & \multicolumn{3}{c}{Validation} & \multicolumn{3}{c}{Test} \\
                             & \multicolumn{1}{c}{P} & \multicolumn{1}{c}{R} & \multicolumn{1}{c}{F1}
                             & \multicolumn{1}{c}{P} & \multicolumn{1}{c}{R} & \multicolumn{1}{c}{F1} \\
      \midrule
      hmm-1	    & 0.6965 & 0.5749 & 0.6299 & 0.6263 & 0.4431 & 0.5190 \\
      hmm-2	    & 0.7047 & 0.6286 & 0.6645 & 0.6480 & 0.5222 & 0.5783 \\
      hmm-3	    & 0.6127 & 0.6141 & 0.6134 & 0.5471 & 0.4634 & 0.5018 \\
      crf	    & 0.7173 & 0.6683 & 0.6920 & 0.6671 & 0.5868 & 0.6244 \\
      lstm-crf	    & 0.8484 & 0.9044 & 0.8755 & 0.8358 & 0.8497 & 0.8427 \\
      lstm-crf-cnn  & 0.9058 & 0.9575 & 0.9309 & 0.8779 & 0.8737 & 0.8758 \\
      lstm-crf-lstm & 0.9134 & 0.9435 & 0.9282 & 0.8920 & 0.8815 & 0.8867 \\
      \bottomrule
    \end{tabular}
  \end{center}
  \caption{Precision, recall and F1 in the complete dataset}
  \label{tab:experiment1}
\end{table}

Experiment 1 consists of testing the models on the complete NER on HTML dataset 
described in section \ref{sec:ner_dataset}. Table \ref{tab:experiment1} shows the 
results for models that employed no features, except for the words in the text
in the case of HMMs and Glove-100 word embeddings in the case of CRFs and Neural Models.
Without carefully designed features and a gazetteer, HMMs and CRFs have a very 
poor performance, achieving only 57.83 for HMM-2 and 62.78 for CRF in the test set. 
The neural models however achieved high F-scores, even with no features. The
plain LSTM-CRF architecture improved significantly in comparison to the conventional CRF,
and the LSTM-CRF architectures with neural character representations boosted the F1 score
by 2.32 for CNN representations and 4.38 for LSTM representations, showing that they are 
able to learn morphological features automatically, and that at least in the NER tasks LSTM 
character representations are superior, perhaps because they are able to differentiate 
suffixes and prefixes, while CNNs are position invariant. This experiment also shows that
pretrained word embeddings can be used to identify named entities in HTML, where
textual context is very limited without resorting to gazetteers.

\begin{table}[h]
  \small
  \begin{center}
    \begin{tabular}{ lllllll }
      \toprule
      \multirow{2}{*}{Model} & \multicolumn{3}{c}{Validation} & \multicolumn{3}{c}{Test} \\
                             & \multicolumn{1}{c}{P} & \multicolumn{1}{c}{R} & \multicolumn{1}{c}{F1}
                             & \multicolumn{1}{c}{P} & \multicolumn{1}{c}{R} & \multicolumn{1}{c}{F1} \\
      \midrule
      hmm-1+F	      & 0.6061 & 0.7282 & 0.6616 & 0.7106 & 0.7633 & 0.7360 \\
      hmm-2+F	      & 0.6279 & 0.7550 & 0.6856 & 0.7521 & 0.7810 & 0.7663 \\
      hmm-3+F	      & 0.6573 & 0.7819 & 0.7142 & 0.7523 & 0.7795 & 0.7657 \\
      hmm-1+F+ST      & 0.7032 & 0.9077 & 0.7925 & 0.7522 & 0.8663 & 0.8052 \\
      hmm-2+F+ST      & 0.7321 & 0.9172 & 0.8143 & 0.7737 & 0.8789 & 0.8230 \\
      hmm-3+F+ST      & 0.7551 & 0.9172 & 0.8283 & 0.7961 & 0.8534 & 0.8237 \\
      crf+F	      & 0.9024 & 0.9049 & 0.9037 & 0.8751 & 0.8227 & 0.8481 \\
      lstm-crf+F      & 0.9430 & 0.9530 & 0.9480 & 0.8998 & 0.8527 & 0.8756 \\
      lstm-crf-cnn+F  & 0.9244 & 0.9715 & 0.9474 & 0.9017 & 0.8973 & 0.8995 \\
      lstm-crf-lstm+F & 0.9465 & 0.9692 & 0.9577 & 0.9108 & 0.8715 & 0.8907 \\
      \bottomrule
    \end{tabular}
  \end{center}
  \caption{Precision, recall and F1 in the complete dataset}
  \label{tab:experiment1f}
\end{table}

Table \ref{tab:experiment1f} shows the results on experiment 1 for models
that employed all features described on Table \ref{tab:features}. In this case
we see that conventional models like HMMs, and CRFs can become competitive with
the right selection of features and a good gazetteer, however they still lose
to the best neural model without features showing their limitation. We can also
see that the self training strategy for HMMs improved the quality of the results
significantly on all cases, boosting both precision and recall.

The neural models also improved a little with the addition of features. The fact that 
the LSTM-CRF gets much closer to the neural models with character representations
shows that the LSTM and CNN character representations were probably able to learn 
at least part of the features automatically, and when we add them explicitly,
the difference in performance becomes less noticeable.

@@

\begin{table}[h]
  \small
  \begin{center}
    \begin{tabular}{ lllllll }
      \toprule
      \multirow{2}{*}{Model} & \multicolumn{2}{c}{Precision} & \multicolumn{2}{c}{Recall} & \multicolumn{2}{c}{F1} \\
                             & \multicolumn{1}{c}{\tiny{M}} & \multicolumn{1}{c}{\tiny{SD}}
			     & \multicolumn{1}{c}{\tiny{M}} & \multicolumn{1}{c}{\tiny{SD}}
			     & \multicolumn{1}{c}{\tiny{M}} & \multicolumn{1}{c}{\tiny{SD}} \\
      \midrule
      hmm-1	      & 0.2856 & 0.1127 & 0.1623 & 0.0810 & 0.2036 & 0.0892 \\
      hmm-2	      & 0.3436 & 0.1134 & 0.2060 & 0.1058 & 0.2483 & 0.0997 \\
      hmm-3	      & 0.2707 & 0.0955 & 0.2163 & 0.0927 & 0.2316 & 0.0793 \\
      crf	      & 0.6249 & 0.0433 & 0.5272 & 0.0348 & 0.5711 & 0.0301 \\
      lstm-crf	      & 0.6982 & 0.0495 & 0.6538 & 0.0532 & 0.6751 & 0.0499 \\
      lstm-crf-cnn    & 0.7906 & 0.0557 & 0.7431 & 0.0687 & 0.7652 & 0.0572 \\
      lstm-crf-lstm   & 0.7935 & 0.0345 & 0.7360 & 0.0424 & 0.7626 & 0.0237 \\
      \bottomrule
    \end{tabular}
  \end{center}
  \caption{Precision, recall and F1 in the small dataset}
  \label{tab:experiment2}
\end{table}

The second experiment aimed at evaluating the performance of sequence models
when the dataset is reduced to only 18 labeled documents. A feasible amount of 
labelling in most WDE tasks. The experiment was repeated in five non overlapping 
sets of training data, and the results were averaged. Table \ref{tab:experiment2} 
describes the results for models without any features. We only present results
for the test set in this experiment. The standard deviation is shown next to each
measure.

We can see that the performance of all models was reduced in comparison 
to experiment 1. That is to be expected.

>>>>

\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{amsfonts}
\usepackage{multirow}

@@

Web data extraction (WDE) methods often rely on hand-coded rules to 
identify and extract data from web pages. These methods 
may be suited to extract information from single websites, but
they usually perform poorly on extraction tasks across different 
websites or different application domains. Statistical methods
are more flexible in this regard. In this paper, we investigate
different approaches to Named Entity Recognition (NER) on HTML, an
essential substask in statistical models of web data extraction.
Webpages can be very different from plain text due to their 
semistructured nature, so we introduced a novel dataset that focuses
on researcher name extraction from faculty listings from numerous
university websites to evaluate different models in the 
NER on HTML task. We found that neural network
architectures with a Conditional Random Fields output layer and
neural character representations are able to achieve an F1-score
of 0.8867 with no inputs besides word embeddings. These NER models 
can be part of a flexible and effective approach to WDE that does not
rely on gazetteers, regular expressions or any hand-coded rules.

@@

@@

HTML documents most often lie in between the structured / unstructured data paradigm.
DOM hierarchy, element disposition, class names, and other features related to 
the document structure and indirectly associated with the data itself can be valuable 
information on the task of identifying entities and determining relationships. Yet we 
cannot expect these features to be completely constrained by an underlying pattern. 
Organization patterns tend to follow some guidelines but are in no way subject to 
strict rules. That is why classical information extraction systems such as automatic
wrapper generators usually do not translate well across different websites.

@@

In recent years, we saw an astonishing progress in Natural Language Processing.
However, despite WDE being a closely related field of research,
many tools rely on heuristics or hand-coded rules to perform record segmentation,
and many methods resort to regular expressions or gazetteer matching to perform attribute
labelling. While these approaches can be sufficient to extract information from webpages
with similar templates and identifying attributes with distinct formats like prices and
dates, we still lack a sufficiently flexible and intelligent approach to handle more complicated cases.
Even when extracting data from web documents only, data records may contain pure text with 
relevant named entities. Also, many WDE algorithms can effectively exploit the semi-
structure of web documents, but when a method relies too much on structural features, it
often produces poor generalization on cross website extraction tasks. Contrastingly, we 
understand that statistical based approaches can be much more flexible.

@@

In 2008, Sarawagi \cite{Sarawagi2008} produced a survey that classified wrappers into
rule-based methods, statistical methods and hybrid models, bringing together 
the fields of named entity recognition, relationship extraction and information extraction. 
The rule based methods encompass most of the 
previous models. The statistical methods convert the extraction task into a token labeling 
task, identifying the target entities through the assignment of labels as in a typical 
Named Entity Recognition task. Traditionally, Hidden Markov Models \cite{Leek1997, Freitag1999}, 
Linear Chain Conditional Random Fields \cite{Lafferty2001}, and Maximum Entropy Taggers 
\cite{McCallum2000} have been the usual choice for linear sequence tagging models.
More recently, with the advancement of Natural Language Processing and Deep Learning, 
neural models outperformed previous NER methods. Huang et. al. \cite{Huang2015} introduced the 
bidirectional Long Short-Term Memory (LSTM) model with a Conditional Random Field (CRF) output layer
for NER. Lample et. al. \cite{Lample2016} incorporated Convolutional Neural Network based character representations 
on top of the architecture. And Ma and Hovy \cite{Ma2016} introduced
LSTM based character representations. Both obtaining improvements with systems that are less reliant
on feature engineering.

@@

Most data extraction systems focus on extracting information from single websites
and are therefore unsuited for cross website extraction tasks. Even unsupervised
approaches that are domain independent, such as RoadRunner \cite{Crescenzi2001} 
and EXALG \cite{Arasu2003} only work well for extracting data from pages generated 
from the same template. A statistical approach to unsupervised domain 
independent web data extraction was described by Zhu et al \cite{Zhu2005}. The 2D CRF 
model takes a web page segmented into data blocks and employs a two dimensional conditional 
random field model to perform attribute labelling. The model was further improved in 
\cite{Zhu2006} to model record segmentation and attribute labelling as a joint task.
Some of the limitations of early unsupervised methods 
were also tackled by ObjectRunner \cite{Abdessalem2010} and AMBER \cite{Furche2012}. 
These methods work by annotating webpages automatically with regular expressions, gazetteers and 
knowledge bases. They can rectify low quality annotations and even improve the annotators
by exploring regular structures in the DOM during the record segmentation phase.

@@

Most web data extraction systems rely on hand crafted rules or gazetteers to perform
attribute annotation. Machine learning approaches to NER can improve annotations of 
more complex entities and even manage entity detection without resorting to a gazetteer.
We explored multiple approaches to the Named Entity Recognition problem in the context 
of a web information extraction task. We first introduce two traditional approaches, 
namely: Hidden Markov Models and Linear Chain Conditional Random Fields. And then
we explore neural network architectures.

@@

A Markov Model is a stochastic model that computes the most probable sequence of states 
given a limited set of observable states $ S = \{s_1, s_2, ..., s_n \} $.
The Hidden Markov Model (HMM) differs from the Markov Model in that it
does not observe the states directly, but rather a probabilistic function of those 
states. For example in NER, the words are observed, however the Named Entity labels
associated with these words are not. Formally, we want to compute the most probable
sequence of labels $ Y = \{y_1, y_2, ..., y_n\} $ for a sequence of observed tokens
$ X = \{x_1, x_2, ..., x_n\} $.

\begin{equation}
Y^* = \argmax_{Y} P(Y|X)
\end{equation}

@@

Since $ P(X) $ is the same for all label sequences $ Y $, we can simply maximize
$ P(X|Y) P(Y) $.

@@

HMM based taggers have been succesfully apllied in many NLP and WDE tasks 
\cite{Rabiner1990, Leek1997, Freitag2000}. They are incredibly fast to train and very 
interpretable, making them a good choice for a first approximation. However, these models 
are highly dependent on the right selection of features, what may outweigh the benefit of a 
small training cost.

@@

A Linear Chain Conditional Random Field (CRF) is the discriminative analog to the HMM,
it was first introduced by Lafferty \cite{Lafferty2001}. It is a distribution $ P(Y|X) $ that takes the form:

\begin{equation}
P(Y|X) = \frac{1}{Z(x)} \prod_{t=1}^{T} exp \left( \sum_{k=1}^{K} \theta_k f_k(y_{t-1}, y_t, X) \right)
\end{equation}
\\

@@

\begin{equation}
Z(x) = \sum_{Y} \prod_{t=1}^{T} exp \left( \sum_{k=1}^{K} \theta_k f_k(y_{t-1}, y_t, x) \right)
\end{equation}
\\

which is a sum over all possible label assignments $ Y $. The partition function can be efficiently
and exactly calculated with the sum-product algorithm. Parameter estimation is usually done through 
negative log likelihood minimization. The function can be optimized with techniques suitable for other 
maximum entropy models such as L-BFGS \cite{Liu1989}. The most likely label sequences can be decoded 
with the Viterbi algorithm, as was the case on HMMs.

@@

\subsection{Neural Network Models}

@@

\begin{figure}
  \centering
	  \includegraphics[width=0.5\textwidth]{pics/cnn}
  \caption{CNN based character representations}
  \label{fig:cnn}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{pics/lstm_char_representations}
  \caption{LSTM based character representations}
  \label{fig:lstm_char}
\end{figure}

@@

\section{NER on HTML Dataset}
\label{sec:ner_dataset}

@@

This task would be similar to the task of labeling authors in comments.
However, we wanted to evaluate models without relying on any sophisticated data record
segmentation system. In many cases, entity annotation may precede the segmentation
phase in web data extraction methods, so annotators have to be able to work with
raw HTML data. Because of the type of documents present in the dataset, many sentences 
contained only one or two tokens, while other documents contained long texts.
Finally, all tokens were tagged using the IOB scheme put forward by
Ramshaw and Marcus \cite{Ramshaw1999}.

@@

\begin{table}[h]
  \small
  \begin{center}
    \begin{tabular}{ lllll }
      \toprule
      Data file & Precision & Recall & F1 & Correct names \\
      \midrule
      Training   & 0.7316 & 0.2303 & 0.3504 & 1341 of 5822 \\ 
      Validation & 0.8474 & 0.2858 & 0.4274 & 511 of 1788 \\ 
      Test       & 0.8717 & 0.3287 & 0.4773 & 890 of 2708 \\ 
      \bottomrule
    \end{tabular}
  \end{center}
  \caption{Gazetteer coverage in each data file}
  \label{tab:gazetteer}
\end{table}

@@

Two experiments were conducted to evaluate the best models for named entity recognition
on HTML in the context of web data extraction. In the first experiment, we trained the NER models
using the entire training set (85 documents). In the second experiment, the training set was 
split into five non overlapping sets with 17 documents, and every model was trained
and tested once using each set, while the test data was kept unchanged. We wanted to verify empirically 
how the models behaved when training data is limited, since labelling NER datasets can be a
very time consuming effort.

Evaluation of model performance in this task was performend with the precision, recall and 
F1 score \cite{Rijsbergen1979}. Precision is the percentage of named entities found by 
the model that are correct. Recall is the percentage of named entities that are present
in the corpus and were found by the model. The F1 score is a composite measure that combines
precision and recall with the formula:

@@

Named entities were only considered to be correct if they were a complete match of the 
corresponding entity in the dataset. The models tested in both experiments are described in Table \ref{tab:models}.
In each experiment, we evaluated variants of each model using the complete set of features described in 
Table \ref{tab:features} and variants that used only lowercase unaccented words for HMMs, 
and only word embeddings in the case of neural models and CRFs. In Tables \ref{tab:experiment1} to 
\ref{tab:experiment2f}, models that used the complete set of features have the suffix "+F". For example,
"lstm-crf+F" is the LSTM-CRF model with all features including GloVe word embeddings, and "lstm-crf" is 
the LSTM-CRF model with only GloVe word embeddings as input.
Additionally, HMMs were also evaluated using the self training strategy. In tables \ref{tab:experiment1f} 
and \ref{tab:experiment2f}, these models have the suffix "+ST".

\subsection{Experiment 1: Complete Dataset}

\begin{table}[h]
  \small
  \begin{center}
    \begin{tabular}{ lllllll }
      \toprule
      \multirow{2}{*}{Model} & \multicolumn{3}{c}{Validation} & \multicolumn{3}{c}{Test} \\
                             & \multicolumn{1}{c}{P} & \multicolumn{1}{c}{R} & \multicolumn{1}{c}{F1}
                             & \multicolumn{1}{c}{P} & \multicolumn{1}{c}{R} & \multicolumn{1}{c}{F1} \\
      \midrule
      hmm-1	    & 0.6965 & 0.5749 & 0.6299 & 0.6263 & 0.4431 & 0.5190 \\
      hmm-2	    & 0.7047 & 0.6286 & 0.6645 & 0.6480 & 0.5222 & 0.5783 \\
      hmm-3	    & 0.6127 & 0.6141 & 0.6134 & 0.5471 & 0.4634 & 0.5018 \\
      crf	    & 0.7173 & 0.6683 & 0.6920 & 0.6671 & 0.5868 & 0.6244 \\
      lstm-crf	    & 0.8484 & 0.9044 & 0.8755 & 0.8358 & 0.8497 & 0.8427 \\
      lstm-crf-cnn  & 0.9058 & 0.9575 & 0.9309 & 0.8779 & 0.8737 & 0.8758 \\
      lstm-crf-lstm & 0.9134 & 0.9435 & 0.9282 & 0.8920 & 0.8815 & 0.8867 \\
      \bottomrule
    \end{tabular}
  \end{center}
  \caption{Precision, recall and F1 in the complete dataset for models using no features}
  \label{tab:experiment1}
\end{table}

Experiment 1 consisted of testing the models in the complete NER on HTML dataset 
described in Section \ref{sec:ner_dataset}. Table \ref{tab:experiment1} shows the 
Precision (P), Recall (R), and F1-scores (F1) for models that employed no features, 
except for lowercase unaccented words
in the case of HMMs and Glove-100 word embeddings in the case of CRFs and Neural Models.

Without carefully designed features and a gazetteer, HMMs and CRFs have a very 
poor performance, achieving an F1-score of only 0.5783 for hmm-2 and 0.6244 for crf
in the test set. This is to be expected, as these models rely on a good feature selection.
The neural models however achieved high F1-scores in the test
data, even with the absence of features. The plain LSTM-CRF architecture improved performance 
significantly in comparison to the conventional CRF, 0.8427 against 0.6244 (a difference 
of 0.2183). Also, the neural character representations boosted performance a little further
reaching an F1-score of 0.8758 for CNN based representations and 0.8867 for LSTM based
representations. The LSTM based representations were superior in modelling
morphological features, perhaps because they are able to differentiate suffixes and prefixes, 
while CNNs filters are position invariant.

The results in experiment 1 also show that pretrained word embeddings can work as a sort of
universal gazetteer. Words with similar embeddings, are likely to belong to the same class.
This knowledge combined with the ability to learn morphological features can make up for 
the scarcity of textual data in some webpages.

\begin{table}[h]
  \small
  \begin{center}
    \begin{tabular}{ lllllll }
      \toprule
      \multirow{2}{*}{Model} & \multicolumn{3}{c}{Validation} & \multicolumn{3}{c}{Test} \\
                             & \multicolumn{1}{c}{P} & \multicolumn{1}{c}{R} & \multicolumn{1}{c}{F1}
                             & \multicolumn{1}{c}{P} & \multicolumn{1}{c}{R} & \multicolumn{1}{c}{F1} \\
      \midrule
      hmm-1+F	      & 0.6061 & 0.7282 & 0.6616 & 0.7106 & 0.7633 & 0.7360 \\
      hmm-2+F	      & 0.6279 & 0.7550 & 0.6856 & 0.7521 & 0.7810 & 0.7663 \\
      hmm-3+F	      & 0.6573 & 0.7819 & 0.7142 & 0.7523 & 0.7795 & 0.7657 \\
      hmm-1+F+ST      & 0.7032 & 0.9077 & 0.7925 & 0.7522 & 0.8663 & 0.8052 \\
      hmm-2+F+ST      & 0.7321 & 0.9172 & 0.8143 & 0.7737 & 0.8789 & 0.8230 \\
      hmm-3+F+ST      & 0.7551 & 0.9172 & 0.8283 & 0.7961 & 0.8534 & 0.8237 \\
      crf+F	      & 0.9024 & 0.9049 & 0.9037 & 0.8751 & 0.8227 & 0.8481 \\
      lstm-crf+F      & 0.9430 & 0.9530 & 0.9480 & 0.8998 & 0.8527 & 0.8756 \\
      lstm-crf-cnn+F  & 0.9244 & 0.9715 & 0.9474 & 0.9017 & 0.8973 & 0.8995 \\
      lstm-crf-lstm+F & 0.9465 & 0.9692 & 0.9577 & 0.9108 & 0.8715 & 0.8907 \\
      \bottomrule
    \end{tabular}
  \end{center}
  \caption{Precision, recall and F1 in the complete dataset for models using all features}
  \label{tab:experiment1f}
\end{table}

Table \ref{tab:experiment1f} shows the results on experiment 1 for models
that employed all the features described in Table \ref{tab:features}.
Conventional models like HMMs, and CRFs can become competitive with
the right selection of features and a good gazetteer, however they still lose
to the best neural model without features, demonstrating their inherent limitations.
We can also see that the self training strategy for HMMs improved the quality of 
the models significantly in all cases, boosting both precision and recall. Finally,
HMMs that employed trigrams or quadrigrams (hmm-2, hmm-3) also performed better in
comparison to regular HMMs.
The neural models also improved a little with the new features. The plain LSTM-CRF
model gets a closer performance to the models that employed neural character 
representations. A fact that hints that the LSTM and CNN character representations 
were able to learn at least part of the morphological features automatically in the
models with no features. So, when these features are added explicitly, the differences 
in performance become less noticeable.

@@

\begin{table}[h]
  \small
  \begin{center}
    \begin{tabular}{ lllllll }
      \toprule
      \multirow{2}{*}{Model} & \multicolumn{2}{c}{Precision} & \multicolumn{2}{c}{Recall} & \multicolumn{2}{c}{F1} \\
                             & \multicolumn{1}{c}{\tiny{M}} & \multicolumn{1}{c}{\tiny{SD}}
			     & \multicolumn{1}{c}{\tiny{M}} & \multicolumn{1}{c}{\tiny{SD}}
			     & \multicolumn{1}{c}{\tiny{M}} & \multicolumn{1}{c}{\tiny{SD}} \\
      \midrule
      hmm-1	      & 0.2856 & 0.1127 & 0.1623 & 0.0810 & 0.2036 & 0.0892 \\
      hmm-2	      & 0.3436 & 0.1134 & 0.2060 & 0.1058 & 0.2483 & 0.0997 \\
      hmm-3	      & 0.2707 & 0.0955 & 0.2163 & 0.0927 & 0.2316 & 0.0793 \\
      crf	      & 0.6249 & 0.0433 & 0.5272 & 0.0348 & 0.5711 & 0.0301 \\
      lstm-crf	      & 0.6982 & 0.0495 & 0.6538 & 0.0532 & 0.6751 & 0.0499 \\
      lstm-crf-cnn    & 0.7906 & 0.0557 & 0.7431 & 0.0687 & 0.7652 & 0.0572 \\
      lstm-crf-lstm   & 0.7935 & 0.0345 & 0.7360 & 0.0424 & 0.7626 & 0.0237 \\
      \bottomrule
    \end{tabular}
  \end{center}
  \caption{Mean Precision, recall and F1 in the test set for models with no features
   trained in five sets with 17 documents}
  \label{tab:experiment2}
\end{table}

The second experiment aimed to evaluate the performance of sequence models when 
the amount of training data is small. The same training data used in Experiment 1 was
split into five non overlapping sets with 17 documents. Each model was trained once 
with each of the five sets and tested with the same test set used in Experiment 1. 
Seventeen documents seem to be a feasible amount of documents to label in most WDE 
tasks. The results for each model was averaged over the five runs.

Table \ref{tab:experiment2} shows the precision, recall and F1-score for the same models
used in Experiment 1, but in this case we only show results for the Test set.
As expected, we can see that the performance of all models with no features was reduced in comparison 
to Experiment 1. HMMs show a terrible performance while CRFs are not 
impacted as much by the reduction in the amount of training examples. Neural models
keep performing better, but they were unable to learn patterns as efficiently in
the smaller training sets. The neural models show a small standard deviation, what
points to the fact that they can retain a consistent performance between different training 
datasets. In this matter, the dropout layers are crucial to produce better generalizations 
and prevent overfitting.
